# Depression-Movement Statistical Tests

This file tests how depression (PHQ-9) relates to head-movement metrics.

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)

set.seed(42)
```

## 1) Load data and build movement metrics

```{r data-load}
data_main <- readr::read_csv("data/data.csv", show_col_types = FALSE) %>%
	clean_names()

head_files <- list.files(
	"data/headtracking-data",
	pattern = "\\.csv$",
	recursive = TRUE,
	full.names = TRUE
)

pick_col <- function(nms, patterns) {
	lower <- tolower(nms)
	exact <- which(lower %in% tolower(patterns))
	if (length(exact) > 0) return(nms[exact[1]])

	fuzzy <- which(stringr::str_detect(lower, paste0("(", paste(patterns, collapse = "|"), ")")))
	if (length(fuzzy) > 0) return(nms[fuzzy[1]])

	NA_character_
}

read_head_file <- function(path) {
	# Some raw CSV fields are mixed-typed; we coerce needed columns manually below.
	df <- suppressWarnings(readr::read_csv(path, show_col_types = FALSE))
	nms <- names(df)

	time_col <- pick_col(nms, c("timestamp", "time", "elapsed_time", "elapsedtime", "seconds", "time_s"))
	rot_change_y_col <- pick_col(nms, c("rotationchangey", "yaw", "head_yaw", "rotation_y", "camerayaw"))
	rot_change_x_col <- pick_col(nms, c("rotationchangex", "pitch", "head_pitch", "rotation_x", "camerapitch"))
	speed_col <- pick_col(nms, c("rotationspeedtotal", "rotation_speed_total", "angular_speed_total"))

	tibble(
		file_name = basename(path),
		video = basename(dirname(path)),
		time_raw = if (!is.na(time_col)) suppressWarnings(as.numeric(df[[time_col]])) else seq_len(nrow(df)),
		rot_change_y = if (!is.na(rot_change_y_col)) suppressWarnings(as.numeric(df[[rot_change_y_col]])) else NA_real_,
		rot_change_x = if (!is.na(rot_change_x_col)) suppressWarnings(as.numeric(df[[rot_change_x_col]])) else NA_real_,
		ang_vel_raw = if (!is.na(speed_col)) suppressWarnings(as.numeric(df[[speed_col]])) else NA_real_
	)
}

head_long <- purrr::map_dfr(head_files, read_head_file) %>%
	group_by(file_name, video) %>%
	mutate(
		time_sec = time_raw - first(time_raw),
		max_time_sec = suppressWarnings(max(time_sec, na.rm = TRUE)),
		time_sec = if_else(is.finite(max_time_sec) & max_time_sec > 2000, time_sec / 1000, time_sec),
		time_sec = if_else(is.na(time_sec), row_number() / 60, time_sec)
	) %>%
	select(-max_time_sec) %>%
	ungroup()

head_summary <- head_long %>%
	group_by(file_name, video) %>%
	arrange(time_sec, .by_group = TRUE) %>%
	mutate(
		dt = pmax(time_sec - lag(time_sec), 1e-6),
		drot_y = rot_change_y - lag(rot_change_y),
		drot_x = rot_change_x - lag(rot_change_x),
		ang_vel_calc = sqrt((drot_y)^2 + (drot_x)^2) / dt,
		ang_vel = dplyr::coalesce(ang_vel_raw, ang_vel_calc)
	) %>%
	summarise(
		rot_change_y_sd = sd(rot_change_y, na.rm = TRUE),
		mean_ang_vel = mean(ang_vel, na.rm = TRUE),
		.groups = "drop"
	)

participant_video_map <- data_main %>%
	select(participant, score_phq, score_gad, score_stai_t, v1, v2, v3, v4, v5) %>%
	pivot_longer(starts_with("v"), names_to = "video", values_to = "file_name") %>%
	mutate(video = factor(video, levels = paste0("v", 1:5)))

participant_head <- head_summary %>%
	mutate(video = factor(video, levels = paste0("v", 1:5))) %>%
	inner_join(participant_video_map, by = c("file_name", "video")) %>%
	group_by(participant, score_phq, score_gad, score_stai_t) %>%
	summarise(
		mean_ang_vel = mean(mean_ang_vel, na.rm = TRUE),
		rot_change_y_sd = mean(rot_change_y_sd, na.rm = TRUE),
		.groups = "drop"
	) %>%
	mutate(
		depression_group = factor(if_else(score_phq >= 10, "Depressed (PHQ>=10)", "Not depressed (PHQ<10)"))
	)

participant_head %>%
	summarise(
		n_participants = n(),
		depressed_n = sum(depression_group == "Depressed (PHQ>=10)", na.rm = TRUE),
		not_depressed_n = sum(depression_group == "Not depressed (PHQ<10)", na.rm = TRUE)
	)
```

## 2) Automatic decision: t-test vs Wilcoxon (by normality)

Rule used:
- If each PHQ binary group has at least 3 observations and both groups pass Shapiro normality (`p > 0.05`), run Welch t-test.
- Otherwise run Wilcoxon rank-sum test.

```{r group-tests}
safe_shapiro_p <- function(x) {
	x <- x[is.finite(x)]
	if (length(x) < 3 || length(x) > 5000) return(NA_real_)
	out <- tryCatch(shapiro.test(x), error = function(e) NULL)
	if (is.null(out)) return(NA_real_)
	out$p.value
}

run_group_test <- function(df, metric_name) {
	d <- df %>%
		select(depression_group, all_of(metric_name)) %>%
		filter(!is.na(depression_group), is.finite(.data[[metric_name]]))

	g_levels <- levels(d$depression_group)
	g1 <- d %>% filter(depression_group == g_levels[1]) %>% pull(all_of(metric_name))
	g2 <- d %>% filter(depression_group == g_levels[2]) %>% pull(all_of(metric_name))

	shapiro_g1 <- safe_shapiro_p(g1)
	shapiro_g2 <- safe_shapiro_p(g2)

	normal_ok <- is.finite(shapiro_g1) && is.finite(shapiro_g2) && shapiro_g1 > 0.05 && shapiro_g2 > 0.05

	if (normal_ok) {
		test_obj <- t.test(as.formula(paste(metric_name, "~ depression_group")), data = d, var.equal = FALSE)
		test_type <- "Welch t-test"
		stat_name <- "t"
		stat_value <- unname(test_obj$statistic)
	} else {
		test_obj <- wilcox.test(as.formula(paste(metric_name, "~ depression_group")), data = d, exact = FALSE)
		test_type <- "Wilcoxon rank-sum"
		stat_name <- "W"
		stat_value <- unname(test_obj$statistic)
	}

	tibble(
		metric = metric_name,
		n_total = nrow(d),
		n_group_1 = length(g1),
		n_group_2 = length(g2),
		shapiro_p_group_1 = shapiro_g1,
		shapiro_p_group_2 = shapiro_g2,
		selected_test = test_type,
		statistic_name = stat_name,
		statistic_value = stat_value,
		p_value = test_obj$p.value
	)
}

group_test_results <- bind_rows(
	run_group_test(participant_head, "mean_ang_vel"),
	run_group_test(participant_head, "rot_change_y_sd")
) %>%
	mutate(p_value_bh = p.adjust(p_value, method = "BH"))

group_test_results
```

## 3) Continuous PHQ association: Pearson vs Spearman

Rule used:
- If both variables are approximately normal (`Shapiro p > 0.05`), use Pearson.
- Otherwise use Spearman.

```{r correlation-tests}
run_corr_test <- function(df, x_name, y_name) {
	d <- df %>%
		select(all_of(c(x_name, y_name))) %>%
		filter(is.finite(.data[[x_name]]), is.finite(.data[[y_name]]))

	shapiro_x <- safe_shapiro_p(d[[x_name]])
	shapiro_y <- safe_shapiro_p(d[[y_name]])

	use_pearson <- is.finite(shapiro_x) && is.finite(shapiro_y) && shapiro_x > 0.05 && shapiro_y > 0.05
	method <- if (use_pearson) "pearson" else "spearman"

	test_obj <- cor.test(d[[x_name]], d[[y_name]], method = method, exact = FALSE)

	tibble(
		outcome_metric = y_name,
		n = nrow(d),
		shapiro_p_phq = shapiro_x,
		shapiro_p_metric = shapiro_y,
		selected_method = str_to_title(method),
		correlation = unname(test_obj$estimate),
		p_value = test_obj$p.value
	)
}

corr_results <- bind_rows(
	run_corr_test(participant_head, "score_phq", "mean_ang_vel"),
	run_corr_test(participant_head, "score_phq", "rot_change_y_sd")
) %>%
	mutate(p_value_bh = p.adjust(p_value, method = "BH"))

corr_results
```

## 4) Optional: collapsed distress group nonparametric check

This keeps PHQ+GAD clinical context (Minimal/Mild/Moderate-Severe).

```{r collapsed-group-kw}
participant_head_collapsed <- participant_head %>%
	mutate(
		distress_collapsed = case_when(
			score_phq >= 10 | score_gad >= 10 ~ "Moderate/Severe",
			score_phq >= 5 | score_gad >= 5 ~ "Mild",
			TRUE ~ "Minimal"
		),
		distress_collapsed = factor(distress_collapsed, levels = c("Minimal", "Mild", "Moderate/Severe"))
	)

kw_results <- tibble(
	metric = c("mean_ang_vel", "rot_change_y_sd"),
	test_obj = list(
		kruskal.test(mean_ang_vel ~ distress_collapsed, data = participant_head_collapsed),
		kruskal.test(rot_change_y_sd ~ distress_collapsed, data = participant_head_collapsed)
	)
) %>%
	mutate(
		chi_sq = map_dbl(test_obj, ~ unname(.x$statistic)),
		df = map_dbl(test_obj, ~ unname(.x$parameter)),
		p_value = map_dbl(test_obj, ~ .x$p.value),
		p_value_bh = p.adjust(p_value, method = "BH")
	) %>%
	select(-test_obj)

kw_results
```

## 5) How to interpret test selection (for your report)

- If `selected_test` is `Welch t-test`, assumptions were acceptable for that metric.
- If `selected_test` is `Wilcoxon rank-sum`, normality was not supported (or sample was too small), so nonparametric was safer.
- If `selected_method` is `Pearson`, both variables were approximately normal.
- If `selected_method` is `Spearman`, rank-based correlation was preferred.

After you run this file, share the printed `group_test_results`, `corr_results`, and `kw_results` tables with me and I will give you precise final wording for your presentation/report.
